---
title: "Useful Functions"
author: "Dave Rosenman"
date: "February 2, 2018"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

What follows is a demonstration of some R functions that I made that (I think) are useful. 

##Combining Columns
The following function, **column_concat** lets you combines multiple columns of strings into a single column. It returns the original columns from the dataframe plus the combined column.

###parameters
- df is the dataframe
-  ... the column names of the columns that will be combined.
- colname is the name of the output column
- sep is the character that will separate the combined column strings

###returns
A dataframe with the first column, "col", being a combination of the column ... separated by "sep" (default ","). Note: If the values in each column for a row are NA, the output for that row in 'col' will be NA. If one or more, but not all, values of the combined columns in a row are NA, all NA's will be remove in 'col' for that row.   

###required packages
- **tidyr**: for the *unite* function 
- **rlang**: for the *enquo* function
- **stringr** for the *string_replace_all* function

```{r}
library(tidyr)
library(rlang)
library(stringr)

column_concat <- function(df,...,col, sep = ",") {
  cols <- quos(...)
  col <- enquo(colname)
  df <- df %>% unite(!!col,!!!cols,sep = sep,remove = FALSE)
  n <- length(cols) 
  na_string <- paste(replicate(n,"NA"), collapse = ",")
  df[,1] <- str_replace_all(df[,1],c(na_string = NA, "NA," = "", ",NA" = ""))
  
  return(df)
}
```
**Example**:

```{r}
df <- data.frame(x = c("U1,U2,U3","U1,U2",NA,NA),
                 y = c("V4,V5",NA,"V8",NA), 
                 z = c('A',NA,'C',NA),
                 stringsAsFactors = FALSE)
print(df)
df_new <- column_concat(df,x,y,z,col = xyz)
print(df_new)
```


