---
title: "Useful Functions"
author: "Dave Rosenman"
date: "February 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

What follows is a demonstration of some R functions that I made that (I think) are useful. 

##Combining Columns
The following function, **column_concat** let's you combine two columns of srins into a single column. It returns the original columns from the dataframe plus the combined column.

###parameters
- df is the dataframe
-  col1 is the name of the first column to be combined
-  col2 is the name of the second column to be combined
- colname is the name of the output column
- sep is the character that will separate the combined column strings

###returns
A dataframe with the first column, 'colname', being a combination of two columns seperated by "sep". If the values in each column for a row are NA, the output for that row in 'colname' will be NA. If one but not both of the values of the combined columns in a row is NA, NA will be removed

###required packages
- **tidyr**: for the *unite* function 
- **rlang**: for the *enquo* function
- **stringr** for the *string_replace_all* function

```{r}
library(tidyr)
library(rlang)
library(stringr)

column_concat <- function(df,col1,col2,colname, sep = ",") {
  c1 <- enquo(col1)
  c2 <- enquo(col2)
  col <- enquo(colname)
  df <- df %>% tidyr::unite(!!col,!!c1,!!c2,sep = sep,remove = FALSE)
  df[,1] <- str_replace_all(df[,1],c("NA,NA" = NA, "NA," = "", ",NA" = ""))
  return(df)
}
```
**Example**:

```{r}
df <- data.frame(x = c("U1,U2,U3","U1,U2","NA","NA"),y = c("V4,V5","NA","V8","NA"), z = 9,stringsAsFactors = FALSE)
print(df)
df_new <- column_concat(df,x,y,xy)
print(df_new)
```
