---
title: "Useful Functions"
author: "Dave Rosenman"
date: "February 2, 2018"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

What follows is a demonstration of some R functions that I made that (I think) are useful. 

## paste_columns
**paste_columns** lets you combines multiple columns of strings into a single column. It returns the original columns from the dataframe plus the combined columns.

### Parameters
- df is the dataframe
-  ... the column names of the columns that will be combined.
- col is the name of the output column
- sep is the character that will separate the combined column strings

### Returns
A dataframe with the first column, 'col', being a combination of the columns '...' separated by "sep" (default ","). Note: If the values in each column for a row are NA, the output for that row in 'col' will be NA. If one or more, but not all, values of the combined columns in a row are NA, all NA's will be remove in 'col' for that row.   

### Required Packages
- **tidyr**: for the *unite* function 
- **rlang**: for the *enquo* function
- **stringr** for the *string_replace_all* function

```{r}
library(tidyr)
library(rlang)
library(stringr)

paste_columns <- function(df,...,col, sep = ",") {
  cols <- quos(...)
  col <- enquo(col)
  df <- df %>% unite(!!col,!!!cols,sep = sep,remove = FALSE)
  n <- length(cols) 
  na_string <- paste(replicate(n,"NA"), collapse = ",")
  df[,1] <- str_replace_all(df[,1],c(na_string = NA, "NA," = "", ",NA" = ""))
  replace(df,df == "NA",NA)
}
```
### Example

```{r}
df <- data.frame(x = c("U1,U2,U3","U1,U2",NA,NA),
                 y = c("V4,V5",NA,"V8",NA), 
                 z = c('A',NA,'C',NA),
                 stringsAsFactors = FALSE)
print(df)
```

```{r}
paste_columns(df,x,y,z,col = xyz)
```

### setexclusive
dplyr contains there set operations functions. All three work on complete rows by comparing the values of each variable a given row. The two input dataframes need to have the same variables. *intersect* returns equivalent rows from the two dataframes, *union* returns unique rows in the two dataframes, and *setdiff* returns rows that are in the first dataframe but not the second. **setexclusive** returns rows that are in either the first dataframe or the second dataframe, but not both dataframes.

### Parameters
- x is the first dataframe
- y is the second dataframe

### Returns
A dataframe consisting of rows in x that are not in dataframe y and rows in y that are not in dataframe x.

### Required Package
- **dplyr**

```{r}
library(dplyr)
setexclusive <- function(x,y) {
  x %>% mutate_if(is.factor,as.character) -> x
  y %>% mutate_if(is.factor,as.character) -> y
  union_df <- union(x,y)
  common_df <- intersect(x,y)
  return(setdiff(union_df,common_df))
  
}
```

### Example
```{r}
library(tibble)
df1 <- tribble(
      ~x,~y,
       1,2,
       2,3,
       3,4,
       4,5
      )
df2 <- tribble(
      ~x,~y,
       1,2,
       9,10,
       2,3
)
#df1 and df2 both have rows (x = 1,y = 2) and (x = 2, y = 3)
#rows exclusive to df1 are rows (x = 3,y = 4) and (x = 4,y = 5)
#the only row exclusive to df2 is (x = 9, y = 10).


setexclusive(df1,df2)

```

